Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-10-05 prop_letter.tex
prop_letter.tex \section{Gleason} Small modification after the meeting..

As an example, consider a three-dimensional Hilbert space with orthonormal
basis $\{\ket{0},\ket{1},\ket{2}\}$ and an observable $\mathbf{O}$
with spectral decomposition $\mathbf{O}=\proj{0}+2\,\proj{1}+3\,\proj{2}$,
i.e., $\lambda_{i}=i$ and $P_{i}=\proj{i-1}$. Two fragments of valid
probability measures~$\mu_{1}$ and $\mu_{2}$ that can be associated
with this space are defined in Table~\ref{tab:quantum-probability-measure}.
\begin{table}
\noindent \centering{}\caption{\label{tab:quantum-probability-measure}Two fragments of valid probability
measures~$\mu_{1}$ and $\mu_{2}$.}
\begin{tabular}{ccc}
\toprule 
\addlinespace
$\ket{\psi}$ & $\mu_{1}(\proj{\psi})$ & $\mu_{2}(\proj{\psi})$\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
$\ket{0}$ & $\frac{1}{2}$ & $\frac{1}{2}$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\ket{1}$ & $\frac{1}{2}$ & $0$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\ket{2}$ & $0$ & $\frac{1}{2}$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\ket{\ps}=\frac{\ket{0}+\ket{1}}{\sqrt{2}}$ & $1$ & $\frac{1}{4}$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\ket{\rmi}=\frac{\ket{0}+\rmi\ket{1}}{\sqrt{2}}$ & $\frac{1}{2}$ & $\frac{1}{4}$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\ket{\ps'}=\frac{\ket{0}+\ket{2}}{\sqrt{2}}$ & $\frac{1}{4}$ & $\frac{1}{2}$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\ket{\rmi'}=\frac{\ket{0}+\rmi\ket{2}}{\sqrt{2}}$ & $\frac{1}{4}$ & $\frac{1}{2}$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\ket{\ps''}=\frac{\ket{1}+\ket{2}}{\sqrt{2}}$ & $\frac{1}{4}$ & $\frac{1}{4}$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\ket{\rmi''}=\frac{\ket{1}+\rmi\ket{2}}{\sqrt{2}}$ & $\frac{1}{4}$ & $\frac{1}{4}$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}
By the Born rule, the first probability measure corresponds to the
quantum system being in the pure state $\ket{\ps}=(\ket{0}+\ket{1})/\sqrt{2}$
and the second corresponds to the quantum system being in the state
$\rho=(\proj{0}+\proj{2})/2$. The expectation values of the observable
$\mathbf{O}$, $\expval{\mathbf{O}}_{\mu_{1,2}}$, are 1.5 in the
first case and 2 in the second.

@@

We will start by reviewing IVPMs in the classical setting and then
propose our quantum generalization. In the classical setting, there
are several proposals for ``imprecise
probabilities''~\citep{Dempster1967,Shafer1976,GilboaSchmeidler1994,Marinacci1999,Weichselberger2000,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
Although these proposals differ in some details, they all share the
fact that the probability $\mu(E)$ of an event~$E$ is generalized from
a single \emph{real number} to an \emph{interval}~$[\ell,r]$
where~$\ell$ intuitively corresponds to the strength of evidence for
the event~$E$ and~$1-r$ corresponds to the strength of evidence
against the same event. More generally, this inverval would be interpreted as
the Guassian width of a probability distribution, but for notational convenience
we will continue in the following to write this as a definite interval.
\yutsung{Check the interpretation in the statistics literature~\citep{HuberRonchetti2009}!
Or we may need to prove it could be interpreted as Guassian width
by ourselves...} The collection of intervals used in the
definition of a probability measure must satisfy certain properties
that we explain next.

@@

\begin{thm}[Non-extensible of Gleason's theorem to general QIVPMs]\label{thm:Non-extensible-of-Gleason's}In
a three-dimensional Hilbert space, there is a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$
such that $\coreBorn\left(\bar{\mu},\events\right)=\emptyset$.\end{thm}

@@

\begin{thm}[Finite-precision Extension of the Born rule]\label{thm:Finite-precision-Born}For
any state $\rho$ on a three-dimensional Hilbert space, a function~$\left[\mu_{\rho}^{\mathrm{B}}\right]_{n}:\events\rightarrow\mathscr{I}_{n}$
defined by $\left[\mu_{\rho}^{\mathrm{B}}\right]_{n}\left(P\right)=\iota_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$
is a QIVPM, called a Born QIVPM.\end{thm}

@@

\begin{thm}[Finite-precision Extension of Gleason's theorem for the
Born QIVPMs]\label{thm:Finite-precision-Gleason}In a three-dimensional
Hilbert space, given any state~$\rho$, if a state $\rho'\in\coreBorn\left(\left[\mu_{\rho}^{\mathrm{B}}\right]_{n},\events\right)$,
then their distance $\left\Vert \rho-\rho'\right\Vert \le\frac{1}{n}$.\end{thm}

>>>>

\begin{comment}
As an example, consider a three-dimensional Hilbert space with orthonormal
basis $\{\ket{0},\ket{1},\ket{2}\}$ and an observable $\mathbf{O}$
with spectral decomposition $\mathbf{O}=\proj{0}+2\,\proj{1}+3\,\proj{2}$,
i.e., $\lambda_{i}=i$ and $P_{i}=\proj{i-1}$. Two fragments of valid
probability measures~$\mu_{1}$ and $\mu_{2}$ that can be associated
with this space are defined in Table~\ref{tab:quantum-probability-measure}.
\begin{table}
\noindent \centering{}\caption{\label{tab:quantum-probability-measure}Two fragments of valid probability
measures~$\mu_{1}$ and $\mu_{2}$.}
\begin{tabular}{ccc}
\toprule 
\addlinespace
$\ket{\psi}$ & $\mu_{1}(\proj{\psi})$ & $\mu_{2}(\proj{\psi})$\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
$\ket{0}$ & $\frac{1}{2}$ & $\frac{1}{2}$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\ket{1}$ & $\frac{1}{2}$ & $0$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\ket{2}$ & $0$ & $\frac{1}{2}$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\ket{\ps}=\frac{\ket{0}+\ket{1}}{\sqrt{2}}$ & $1$ & $\frac{1}{4}$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\ket{\rmi}=\frac{\ket{0}+\rmi\ket{1}}{\sqrt{2}}$ & $\frac{1}{2}$ & $\frac{1}{4}$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\ket{\ps'}=\frac{\ket{0}+\ket{2}}{\sqrt{2}}$ & $\frac{1}{4}$ & $\frac{1}{2}$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\ket{\rmi'}=\frac{\ket{0}+\rmi\ket{2}}{\sqrt{2}}$ & $\frac{1}{4}$ & $\frac{1}{2}$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\ket{\ps''}=\frac{\ket{1}+\ket{2}}{\sqrt{2}}$ & $\frac{1}{4}$ & $\frac{1}{4}$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\ket{\rmi''}=\frac{\ket{1}+\rmi\ket{2}}{\sqrt{2}}$ & $\frac{1}{4}$ & $\frac{1}{4}$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}
By the Born rule, the first probability measure corresponds to the
quantum system being in the pure state $\ket{\ps}=(\ket{0}+\ket{1})/\sqrt{2}$
and the second corresponds to the quantum system being in the state
$\rho=(\proj{0}+\proj{2})/2$. The expectation values of the observable
$\mathbf{O}$, $\expval{\mathbf{O}}_{\mu_{1,2}}$, are 1.5 in the
first case and 2 in the second.
\end{comment}

@@

We will start by reviewing IVPMs in the classical setting and then
propose our quantum generalization. In the classical setting, there
are several proposals for ``imprecise
probabilities''~\citep{Dempster1967,Shafer1976,GilboaSchmeidler1994,Marinacci1999,Weichselberger2000,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
Although these proposals differ in some details, they all share the
fact that the probability $\mu(E)$ of an event~$E$ is generalized from
a single \emph{real number} to an \emph{interval}~$[\ell,r]$
where~$\ell$ intuitively corresponds to the strength of evidence for
the event~$E$ and~$1-r$ corresponds to the strength of evidence
against the same event. More generally, this inverval would be interpreted as
the Gaussian width of a probability distribution, but for notational
convenience we will continue in the following to write this as a definite interval.
\yutsung{Check the interpretation in the statistics literature~\citep{HuberRonchetti2009}!
Or we may need to prove it could be interpreted as Guassian width
by ourselves...} The collection of intervals used in the
definition of a probability measure must satisfy certain properties
that we explain next.

@@

\begin{thm}[Empty core for general QIVPMs]\label{thm:Non-extensible-of-Gleason's}In
a three-dimensional Hilbert space, there exists a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$
such that $\coreBorn\left(\bar{\mu},\events\right)=\emptyset$.\end{thm}

@@

\begin{thm}[A finite-precision Extension of the Born rule]\label{thm:Finite-precision-Born}For
any state $\rho$ on a three-dimensional Hilbert space, a function~$\left[\mu_{\rho}^{\mathrm{B}}\right]_{n}:\events\rightarrow\mathscr{I}_{n}$
defined by $\left[\mu_{\rho}^{\mathrm{B}}\right]_{n}\left(P\right)=\iota_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$
is a QIVPM, called a Born QIVPM.\end{thm}

@@

\begin{thm}[A finite-precision extension of weak Gleason's theorem
for Born QIVPMs]\label{thm:Finite-precision-Gleason}In a three-dimensional
Hilbert space, given any state~$\rho$, if a state $\rho'\in\coreBorn\left(\left[\mu_{\rho}^{\mathrm{B}}\right]_{n},\events\right)$,
then their distance $\left\Vert \rho-\rho'\right\Vert \le\frac{1}{n}$.\end{thm}
